toyData

### 2017-10-31 latex/article_vol3.Rnw
Convert manuscript from tex -> Rnw

@@

\begin{Schunk}
\begin{Sinput}
R> library("dataMaid")
R> data("toyData")
R> toyData
\end{Sinput}
\begin{Soutput}
   var1 var2  var3       var4 var5       var6
1   red    1     a -0.6264538    1 Irrelevant
2   red    1     a  0.1836433    2 Irrelevant
3   red    1     a -0.8356286    3 Irrelevant
4   red    2     a  1.5952808    4 Irrelevant
5   red    2     a  0.3295078    5 Irrelevant
6   red    6     b -0.8204684    6 Irrelevant
7   red    6     b  0.4874291    7 Irrelevant
8   red    6     b  0.7383247    8 Irrelevant
9   red  999     c  0.5757814    9 Irrelevant
10  red   NA     c -0.3053884   10 Irrelevant
11 blue    4     c  1.5117812   11 Irrelevant
12 blue   82     .  0.3898432   12 Irrelevant
13 blue   NA       -0.6212406   13 Irrelevant
14 <NA>  NaN other -2.2146999   14 Irrelevant
15 <NA>    5 OTHER  1.1249309   15 Irrelevant
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, output = "html", onlyProblematic = TRUE,
+    maxProbVals = 2, replace = TRUE)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> allSummaryFunctions()
\end{Sinput}
\begin{Soutput}
----------------------------------------------------------------------------
name           description                     classes                      
-------------- ------------------------------- -----------------------------
centralValue   Compute median for numeric      character, Date, factor,     
               variables, mode for             integer, labelled, logical,  
               categorical variables           numeric

countMissing   Compute proportion of missing   character, Date, factor,     
               observations                    integer, labelled, logical,  
                                               numeric

minMax         Find minimum and maximum        integer, numeric, Date       
               values

quartiles      Compute 1st and 3rd quartiles   Date, integer, numeric

uniqueValues   Count number of unique values   character, Date, factor,     
                                               integer, labelled, logical,  
                                               numeric

variableType   Data class of variable          character, Date, factor,     
                                               integer, labelled, logical,  
                                               numeric                      
----------------------------------------------------------------------------
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> setSummaries()
\end{Sinput}
\begin{Soutput}
$character
[1] "variableType" "countMissing" "uniqueValues" "centralValue"

$factor
[1] "variableType" "countMissing" "uniqueValues" "centralValue"

$labelled
[1] "variableType" "countMissing" "uniqueValues" "centralValue"

$numeric
[1] "variableType" "countMissing" "uniqueValues" "centralValue" "quartiles"   
[6] "minMax"

$integer
[1] "variableType" "countMissing" "uniqueValues" "centralValue" "quartiles"   
[6] "minMax"

$logical
[1] "variableType" "countMissing" "uniqueValues" "centralValue"

$Date
[1] "variableType" "countMissing" "uniqueValues" "centralValue" "minMax"      
[6] "quartiles" 
\end{Soutput}
\end{Schunk} %$

@@

\begin{Schunk}
\begin{Sinput}
R> defaultCaracterSummaries()
\end{Sinput}
\begin{Soutput}
[1] "variableType" "countMissing" "uniqueValues" "centralValue"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, 
+    summaries = setSummaries(numeric = c("variableType","centralValue),
+    integer = c("variableType", "centralValue")))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, 
+    summaries = setSummaries(all = c("variableType", "centralValue")))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> defaultFactorChecks()
\end{Sinput}
\begin{Soutput}
[1] "identifyMissing"    "identifyWhitespace" "identifyLoners"    
[4] "identifyCaseIssues" "identifyNums"      
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, 
+    checks = setChecks(factor = "identifyWhitespace"))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, checks = setChecks(factor = NULL))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
>R complexData <- data.frame(complexVar = complex(100, real = 1:100, 
+    imaginary = 3), numericVar = 1:100))
>R makeDataReport(complexData, treatXasY(complex = "numeric"))
\end{Sinput}
\end{Schunk}

In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a \code{complex} variable as a \code{numeric} means dropping the imaginary part of the complex numbers - which was the only thing setting the two variables apart in the first place.

@@

\begin{Schunk}
\begin{Sinput}
R> check(toyData$var2)
\end{Sinput}
\begin{Soutput}
$identifyMissing
The following suspected missing value codes enter as regular values: 999, NaN.
$identifyOutliers
Note that the following possible outlier values were detected: 82, 999.
\end{Soutput}
\end{Schunk}%$

@@

\begin{Schunk}
\begin{Sinput}
R> check(toyData$var2, checks = setChecks(numeric = "identifyMissing"))
\end{Sinput}
\begin{Soutput}
$identifyMissing
The following suspected missing value codes enter as regular values: 999, NaN.
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> identifyMissing(toyData$var2)
\end{Sinput}
\begin{Soutput}
The following suspected missing value codes enter as regular values: 999, NaN.
\end{Soutput}
\end{Schunk} %$

@@

\begin{Schunk}
\begin{Sinput}
R> missVar2 <- identifyMissing(toyData$var2)
R> str(missVar2)
\end{Sinput}
\begin{Soutput}
List of 3
 $ problem      : logi TRUE
 $ message      : chr "The following suspected missing value codes
    enter as regular values: \\\"999\\\", \\\"NaN\\\"."
 $ problemValues: num [1:2] 999 NaN
 - attr(*, "class")= chr "checkResult"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
R> identifyMissing(toyData$var2)
\end{Sinput}
\begin{Soutput}
No problems found.
\end{Soutput}
\end{Schunk} %$

@@

\begin{Schunk}
\begin{Sinput}
R> visualize(toyData$var2)
R> summarize(toyData$var2)
\end{Sinput}
\begin{Soutput}
     Feature                   Result       
[1,] "Variable type"           "numeric"    
[2,] "Number of missing obs."  "4 (26.67 %)"
[3,] "Number of unique values" "6"          
[4,] "Median"                  "4"          
[5,] "1st and 3rd quartiles"   "1.5; 6"     
[6,] "Min. and max."           "1; 82"      
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> summarize(toyData$var2, 
+    summaries = setSummaries(
+    numeric = defaultNumericSummaries(remove = c("variableType", 
+    "countMissing"))))
\end{Sinput}
\begin{Soutput}
     Feature         Result 
[1,] "Median"        "4"    
[2,] "Min. and max." "1; 82"
\end{Soutput}
\end{Schunk} %$

@@

\begin{Schunk}
\begin{Sinput}
R> data(presidentData)
R> makeDataReport(presidentData, replace = TRUE, 
+    treatXasY = list("Name" = "character"),
+    checks = setChecks(character = 
+    defaultCharacterChecks(remove = "identifyLoners")),
+    reportTitle = "Dirty president data")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> presidentData$lastName[presidentData$lastName == " Truman"] <- 
+    "Truman"
R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- 
+    "New York"
R> presidentData$assassinationAttempt <- 
+    factor(presidentData$assassinationAttempt)
R> presidentData$ageAtInauguration <- 
+    as.numeric(presidentData$ageAtInauguration)
\end{Sinput}
\end{Schunk}%$

@@

\begin{Schunk}
\begin{Sinput}
>R presidentData$ageAtInauguration <- 
+    as.numeric(as.character(presidentData$ageAtInauguration))
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> presidentData$lastName[presidentData$firstName == "."]
\end{Sinput}
\begin{Soutput}
[1] "Trump"
\end{Soutput}
\begin{Sinput}
R> presidentData$firstName[presidentData$firstName == "."] <- "Donald"
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> birthdayOutlierVal <- 
+    identifyOutliers(presidentData$birthday)$problemValues
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> presidentData[presidentData$birthday == birthdayOutlierVal, ]
\end{Sinput}
\begin{Soutput}
      lastName firstName orderOfPresidency   birthday stateOfBirth
46 Arathornson   Aragorn                 0 1300-03-01       Gondor
   assassinationAttempt  sex ethnicity precidencyYears ageAtInauguration
46                    1 Male Caucasian              NA              88
   favoriteNumber
46           8+0i
\end{Soutput}
\end{Schunk}%$

@@

\begin{Schunk} 
\begin{Sinput}
R> presidentData <- 
+    presidentData[presidentData$birthday != birthdayOutlierVal, ]
\end{Sinput}
\end{Schunk} %$

@@

\begin{Schunk}
\begin{Sinput}
>R summarize(presidentData$presidencyYears)
\end{Sinput}
\begin{Soutput}
$variableType
Variable type: numeric
$countMissing
Number of missing obs.: 1 (2.22 %)
$uniqueValues
Number of unique values: 11
$centralValue
Median: 4
$quartiles
1st and 3rd quartiles: 4; 8
$minMax
Min. and max.: 0; Inf
\end{Soutput}
\end{Schunk} %$

@@

\begin{Schunk}
\begin{Sinput}
R> presidentData[is.na(presidentData$presidencyYears) | 
+     presidentData$presidencyYears %in% 
+     identifyOutliers(presidentData$presidencyYears)$problemValues, 
+     c("firstName", "lastName", "presidencyYears")]
\end{Sinput}
\begin{Soutput}
   firstName lastName presidencyYears
9    William Harrison               0
12   Zachary   Taylor               1
20     James Garfield               0
44    Barack    Obama             Inf
45    Donald    Trump              NA
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> presidentData$presidencyYears[presidentData$lastName == "Obama"] <- 8
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
makeDataReport(presidentData, vol = "_cleaned", 
+    treatXasY = list(Name = "character", complex = "numeric"),
+    checks = setChecks(character = 
+    defaultCharacterChecks(remove = "identifyLoners")),
+    reportTitle = "Dirty president data - cleaned")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, onlyProblematic = TRUE, mode = "check")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> toyChecks <- check(toyData)
R> foo <- function(x) {
+    any(sapply(x, function(y) y[["problem"]]))
+  }
R> sapply(toyChecks, foo)
\end{Sinput}
\begin{Soutput}
 var1  var2  var3  var4  var5  var6 
 TRUE  TRUE  TRUE  TRUE  TRUE FALSE 
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, standAlone = FALSE)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, standAlone = FALSE, output = "html", 
+    twoCols = FALSE)
\end{Sinput}
\end{Schunk}

@@

Alternatively, one can create the usual output report, not render it and then manually edit the produced \proglang{R} markdown file as wished. This can be done with the following command:
\begin{Schunk}
\begin{Sinput}
R> makeDataReport(toyData, render = FALSE, openResult = FALSE)
\end{Sinput}
\end{Schunk}
After editing, the file can be rendered by calling the \code{render} function:
\begin{Schunk}
\begin{Sinput}
R> render("dataMaid_toyData.Rmd", quiet = FALSE)
\end{Sinput}
\end{Schunk}

>>>>

<<echo=FALSE>>=
# options(prompt = "R> ", width=77, continue = "+  ")
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
@

@@

<<>>=
library("dataMaid")
data("toyData")
toyData
@

<<>>=
makeDataReport(toyData)
@

@@

<<>>=
makeDataReport(toyData, output = "html", onlyProblematic = TRUE,
   maxProbVals = 2, replace = TRUE)
@

@@

<<>>=
makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE, replace=TRUE)
@

@@

<<>>=
allSummaryFunctions()
@

@@

<<>>=
setSummaries()
@

@@

<<>>=
defaultCharacterSummaries()
@

@@

<<>>=
makeDataReport(toyData, replace=TRUE,
  summaries = setSummaries(numeric = c("variableType","centralValue"),
  integer = c("variableType", "centralValue")))
@

@@

<<>>=
makeDataReport(toyData, replace=TRUE,                               
  summaries = setSummaries(all = c("variableType", "centralValue")))
@

@@

<<>>=
defaultFactorChecks()
@

@@

<<>>=
makeDataReport(toyData, replace=TRUE,
  checks = setChecks(factor = "identifyWhitespace"))
@

@@

<<>>=
makeDataReport(toyData, checks = setChecks(factor = NULL), replace=TRUE)
@

@@

<<>>=
makeDataReport(toyData, preChecks = c("isKey", "isSupported"), replace=TRUE)
@

@@

<<>>=
makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), replace=TRUE)
@

@@

<<warning=FALSE>>=
complexData <- data.frame(complexVar = complex(100, real = 1:100, imaginary = 3), numericVar = 1:100)
makeDataReport(complexData, treatXasY=list(complex = "numeric"), replace=TRUE)
@

In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a \code{complex} variable as a \code{numeric} means dropping the imaginary part of the complex numbers which was the only thing setting the two variables apart in the first place.

@@

<<>>=
check(toyData$var2)
@

@@

<<>>=
check(toyData$var2, checks = setChecks(numeric = "identifyMissing"))
@

@@

<<>>=
identifyMissing(toyData$var2)
@

@@

<<>>=
missVar2 <- identifyMissing(toyData$var2)
str(missVar2)
@

@@

<<>>=
toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
identifyMissing(toyData$var2)
@

@@

<<>>=
visualize(toyData$var2)
summarize(toyData$var2)
@

@@

<<>>=
summarize(toyData$var2, 
  summaries = setSummaries(
  numeric = defaultNumericSummaries(remove = c("variableType", 
  "countMissing"))))
@

@@

<<>>=
data(presidentData)
makeDataReport(presidentData, replace = TRUE, 
  treatXasY = list("Name" = "character"),
  checks = setChecks(character = 
  defaultCharacterChecks(remove = "identifyLoners")),
  reportTitle = "Dirty president data")
@

@@

<<>>=
presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- "New York"
presidentData$assassinationAttempt <- factor(presidentData$assassinationAttempt)
presidentData$ageAtInauguration <- as.numeric(presidentData$ageAtInauguration)
@

@@

<<>>=
presidentData$ageAtInauguration <- as.numeric(as.character(presidentData$ageAtInauguration))
@

@@

<<>>=
presidentData$lastName[presidentData$firstName == "."]
@

<<>>=
presidentData$firstName[presidentData$firstName == "."] <- "Donald"
@

@@

<<>>=
birthdayOutlierVal <- 
+    identifyOutliers(presidentData$birthday)$problemValues
@

@@

<<>>=
presidentData[presidentData$birthday == birthdayOutlierVal, ]
@

@@

<<>>=
presidentData <- 
  presidentData[presidentData$birthday != birthdayOutlierVal, ]
@

@@

<<>>=
summarize(presidentData$presidencyYears)
@

@@

<<>>=
presidentData[is.na(presidentData$presidencyYears) | 
   presidentData$presidencyYears %in% 
   identifyOutliers(presidentData$presidencyYears)$problemValues, 
   c("firstName", "lastName", "presidencyYears")]
@

@@

<<>>=
presidentData$presidencyYears[presidentData$lastName == "Obama"] <- 8
@

@@

<<>>=
makeDataReport(presidentData, vol = "_cleaned", 
  treatXasY = list(Name = "character", complex = "numeric"),
  checks = setChecks(character = 
  defaultCharacterChecks(remove = "identifyLoners")),
  reportTitle = "Dirty president data - cleaned",
  replace=TRUE)
@

@@

<<>>=
makeDataReport(toyData, onlyProblematic = TRUE, mode = "check", replace=TRUE)
@

@@

<<>>=
toyChecks <- check(toyData)
foo <- function(x) {
  any(sapply(x, function(y) y[["problem"]]))
}
sapply(toyChecks, foo)
@

@@

<<eval=FALSE>>=
makeDataReport(toyData, standAlone = FALSE)
@

@@

<<eval=FALSE>>=
makeDataReport(toyData, standAlone = FALSE, output = "html", 
  twoCols = FALSE)
@

@@

Alternatively, one can create the usual output report, not render it and then manually edit the produced \proglang{R} markdown file as wished. This can be done with the following command:

<<eval=FALSE>>=
makeDataReport(toyData, render = FALSE, openResult = FALSE)
@

After editing, the file can be rendered by calling the \code{render} function:

<<eval=FALSE>>=
render("dataMaid_toyData.Rmd", quiet = FALSE)
@

### 2017-10-31 latex/dataMaid_toyData.Rmd
Convert manuscript from tex -> Rnw

---
dataMaid: yes
title: toyData
subtitle: "Autogenerated data summary from dataMaid"
date: "`r Sys.Date()`"
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
# Data cleaning summary
The dataset examined has the following dimensions:

------------------------------
Feature                 Result
--------------------- --------
Number of rows              15

Number of variables          6
------------------------------

@@

# Variable list
## **var5**

@@

## **var6**

@@

This report was created by dataMaid v0.9.2.

>>>>

---
dataMaid: yes
title: toyData
subtitle: "Autogenerated data summary from dataMaid"
date: 2017-10-31 15:34:55
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations         15

Number of variables             6
---------------------------------

@@

# Summary table

----------------------------------------------------------------------------------
&nbsp;   Variable class     # unique values   Missing observations  Any problems? 
-------- ---------------- ----------------- ---------------------- ---------------
[var1]   factor                           3                13.33 %    $\times$

[var2]   numeric                          7                26.67 %    $\times$

[var3]   factor                           7                 0.00 %    $\times$

[var4]   numeric                         15                 0.00 %    $\times$

[var5]   factor                          15                 0.00 %    $\times$

[var6]   factor                           1                 0.00 %    $\times$    
----------------------------------------------------------------------------------

# Variable list
## var1

- Note that the following levels have at most five observations: \"blue\".

## var2

- Note that the following possible outlier values were detected: \"82\".

## var3

- The following suspected missing value codes enter as regular values: \" \", \".\".

- The following values appear with prefixed or suffixed white space: \" \".

- Note that the following levels have at most five observations: \" \", \".\", \"OTHER\", \"a\", \"b\", \"c\", \"other\".

- Note that there might be case problems with the following levels: \"OTHER\", \"other\".

## var4

- Note that the following possible outlier values were detected: \"1.12\", \"1.51\", \"1.6\".

## var5

@@

## var6

@@

Report generation information:

 *  Created by Claus Thorn Ekstrøm.

 *  Report creation time: Tue Oct 31 2017 15:34:55

 *  dataMaid v0.9.7.9000 [Pkg: 2017-10-31 from local (ekstroem/dataMaid@NA)]

 *  R version 3.4.1 (2017-06-30).

 *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS Sierra 10.12.6).

 *  Function call: `makeDataReport(data = toyData, onlyProblematic = TRUE, mode = "check", 
    replace = TRUE)`
